Minimal trap-spaces of Logical models are maximal siphons of their Petri-net encoding

### 2022-05-01 cmsb22.tex
Comments, title
Getting some text in

\documentclass[runningheads]{llncs}
\usepackage[T5]{fontenc}
% \usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}

@@

@@

@@

\subsubsection{Petri-net encoding of Logical models}

Original encoding~\cite{chaouiya2004qualitative,chaouiya2011petri} vs. more recent one~\cite{chatain2014characterization}.
All identical on Boolean models?

@@

@@

SBML-Qual~\cite{chaouiya2013sbml}/BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} to PNML\footnote{\url{https://www.pnml.org/}} works via GINsim~\cite{chaouiya2012logical}/BioLQM\footnote{http://www.colomoto.org/biolqm/} with
\begin{verbatim}
java -jar GINsim.jar -lqm <input.sbml/input.bnet> <output.pnml>
\end{verbatim}

@@

>>>>

\documentclass[runningheads]{llncs}
\usepackage[T5]{fontenc}
% \usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\renewcommand\UrlFont{\color{blue}\rmfamily}

\newcommand{\vangiang}[1]{\textcolor{magenta}{#1}}
\newcommand{\sylvain}[1]{\textcolor{teal}{#1}}

@@

\sylvain{We could go for a completely different title, e.g. focused on performance much more, like ``Scaling-up attractor computation for Logical models''}

\cite{glass1973logical,thomas1973boolean,thomas1990biological}

@@

Purely boolean, asynchronous.

@@

\subsubsection{Petri-net encoding of Logical models}
\label{sec:encoding}

The link between Logical models \emph{à la} Thomas and Petri-nets was originally established in~\cite{chaouiya2004qualitative} in order to make available formal methods like model-checking for the analysis of such systems.
The basic encoding only holds for purely Boolean models but was later extended to multi-valued models in two ways, either in \cite{chaouiya2011petri} with non 1-safe Petri-nets or more recently in~\cite{chatain2014characterization} with 1-safe nets but many more places.

Since our study is focused on Boolean models, we will only briefly recall the original encoding here.
Its basis is that every \emph{gene} of the original model is represented by two separate places, corresponding to its two states, active, and inactive.
Each conjunct of the logical function that activate the \emph{gene} will lead to a transition, consuming the inactive place, producing the active place, and with all other literals both consumed and produced.
And conversely for the inactivation.

Note that given a Logical model in the standard SBML-Qual format~\cite{chaouiya2013sbml}, i.e., one of the packages of SBML v3~\cite{keating2020sbml}, one can easily obtain its Petri-net encoding in the PNML\footnote{\url{https://www.pnml.org/}} standard using the BioLQM\footnote{\url{http://www.colomoto.org/biolqm/}} library.
This piece of software extracted from GINsim~\cite{chaouiya2012logical} and part of the CoLoMoTo\footnote{\url{http://colomoto.org/}} software suite allows for easy conversion between standard formats.
It also accepts many other common formats for Logical models, notably the \verb|.bnet| files of the  BoolNet~\cite{mussel2010boolnet,klarner2017pyboolnet} tools.

Conversion is executed as follows:

\noindent{\small \verb|java -jar GINsim.jar -lqm <input.{sbml,bnet,zginml,...}> <output.pnml>|}

@@

Siphons are a classical property of place-transition nets~\cite{peterson81petri}.
Note however that the use of siphons for the analysis of biological models, though it is not new~\cite{angeli2007petri,angeli2011persistence,degrand2020graphical}, has been mostly relevant to the ODE-based continuous semantics of Chemical Reaction Networks.

We recall here the basic definition:

\begin{definition}

  A siphon is a set of places such that for each edge from a transition to any place of the siphon, there is an edge from a place of the siphon to that transition.

\end{definition}

Intuitively a siphon is a set of places that once empty remains empty, i.e.,
a set of species that cannot be produced again once they have been completely
consumed.

@@

@@

Using BioLQM~\ref{sec:encoding}
